---
title: "Homework 2"
author: "Teshawna Badu"
date: 10/09/2021
output: github_document
---

I'm an R Markdown document!

```{r setup}
library(tidyverse)
library(readxl)
library(readr)
```

Download the trashwheel dataset 
```{r}
trashwheel_df = 
  readxl::read_excel("TrashWheelCollection.xlsx",
  sheet = "MrTrashWheel", range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  mutate(sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)) %>% 
  drop_na(dumpster)
```

# Problem 1: 2018 and 2019 precipitation data 
```{r}
precip_18 = 
  readxl::read_excel("TrashWheelCollection.xlsx",
  sheet = "2018 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>%
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)


precip_19 = 
  readxl::read_excel("TrashWheelCollection.xlsx",
  sheet = "2019 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>%
  drop_na(month) %>% 
  mutate(year = 2019) %>% 
  relocate(year)
```

# Combining annual precipitation 
x
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

Precipitation_df = 
  bind_rows(precip_19, precip_18)

left_join(Precipitation_df, month_df, by = "month")

summary(Precipitation_df)
```

The Mr. Trash Wheel is a water wheel trash collector located in Inner Harbor in Baltimore, Maryland.  The data in this dataset includes `r colnames(Precipitation_df)`. The size of dataset we have is `r nrow(Precipitation_df) * ncol(Precipitation_df)`.The total precipitation in 2018 is
`r sum(Precipitation_df[Precipitation_df["year"] == 2018, "total"])`. The median number of sports balls in a dumpster in 2019 is `r trashwheel_df %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`.


#  Problem 2
# clean pols-month
```{r}
pols_df = read_csv("./FiveThirtyEight/pols-month.csv") %>% 
 janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), convert = T) %>% 
  mutate(month = month.name[month],
         president = recode(prez_gop, `0` = "dem", `1` = "gop", `2` = "gop")) %>%
  select(-day, -starts_with("prez"))
```

# clean snp
```{r}
snp_path = "./FiveThirtyEight/snp.csv"

my_parse_date = function(date){
  date = parse_date(date, format = '%m/%d/%y')
  date[format(date, "%y") >= 50] = parse_date(format(date[format(date, "%y") >= 50], "19%y-%m-%d"))
  return(date)
}
snp_df =
read_csv(snp_path) %>% 
janitor::clean_names() %>% 
mutate(date = my_parse_date(date)) %>% 
separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
arrange(year, month) %>% 
mutate(month = month.name[month]) %>% 
relocate(year, month) %>% 
select(-day)
```


# clean unemployment
```{r}
unemploy_path = "./FiveThirtyEight/unemployment.csv"
month_dict = month.name
names(month_dict) = month.abb

unemploy_df = 
  read_csv(unemploy_path) %>%
  pivot_longer(cols = "Jan":"Dec", names_to = "month", values_to = "unemployment_rate") %>% 
  mutate(month = as.character(month_dict[month])) %>% 
  janitor::clean_names()
```


# Merge 

```{r}
Fivethirty_df = 
  left_join(pols_df, snp_df) %>% 
  left_join(unemploy_df)
```
The pols-month is a dataset related to the number of national politicians who are democratic or republican at any given time. The `pols_df` dataset contains `r ncol(pols_df)` variables and `r nrow(pols_df)` observations. The key variables include`r colnames(pols_df)[3:9]`, from 
`r min(pull(pols_df, year))` to `r max(pull(pols_df, year))`.


Snp is a dataset related to Standard & Poor’s stock market index (S&P). It contains `r ncol(snp_df)` variables and `r nrow(snp_df)` observations.The key variables include `r colnames(snp_df)[3]`, from `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`.


The unemployment dataset consists of information related to unemployment percentage with respect to years and months. It contains `r ncol(unemploy_df)` variables and `r nrow(unemploy_df)` 
observations.The key variables include `r ncol(unemploy_df)` variables and `r nrow(unemploy_df)` 
observations. It provides relevant information of unemployment rate per month in US by the variable`r colnames(unemploy_df)[3]` from `r min(pull(unemploy_df, year))` to `r max(pull(unemploy_df, year))`.

